Empirical Analysis of the Role of Energy in Economic Growth

### 2013-01-31 EconGrowth.Rnw
Now using global variables for much of the control parameters for the graphs. This will ensure consistent appearance among all lattice (9-panel) graphs.

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countries <- c(US="US", UK="UK", JP="JP", CN="CN", ZA = "ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
# List of energy types
energyTypes <- c(Q = "Q", X = "X", U = "U")
@

@@

<<Country Factors Graph, echo=FALSE>>=
createCountryFactorsGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graphType <- "l"
  lineTypes <- c(1, 5, 2, 4, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(1, 1, 1, 1, 1) #line widths
  colors <- c("black", "black", "black", "black", "black") #line and point colors
  lineSpec <- list(lty=lineTypes, lwd=lineWidths, col=colors)
  graph <- xyplot(iCapStk+iLabor+iQ+iX+iU ~ Year, data=dataTable,
                  type=graphType,
                  par.settings = list(superpose.line = lineSpec),
                  key=list(text=list(c("k", "l", "q", "x", "u")),
                           type=graphType,
                           lines=lineSpec,
                           columns=1, x=0.0, y=0.98),
                  scales=list(cex=1.0, #controls text size on scales
                              tck=-0.5), #controls tick mark length
                  ylab="Indexed (1980=1)")
  return(graph)
}
@
<<Factors_Lattice_Graph, fig.width=6.5, fig.height=7.5, fig.cap="GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", echo=FALSE>>=
createFactorsLatticeGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graphType <- "b"
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "black") #line and point colors
  symbols <- c(1, NA_integer_, NA_integer_, NA_integer_, NA_integer_, NA_integer_)  #NA_integer_ gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(c(7,6,3,1,8,5,4,2,9)), #orders the panels. Default is alphabetical.
                  layout=c(3,3), #indicates a 3x3 arrangement of panels.
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=0.85, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=1, x=0.01, y=0.95), #We want only one column and sets location of the key
                  scales=list(cex=0.75, #controls text size on scales.
                              tck=-0.5, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph("All") #Reads data from the "AllData.txt" file.
@

@@

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds to the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(){
  preds <- lapply(countries, cobbDouglasPredictions)
  out <- rbind(preds[["US"]], preds[["UK"]], preds[["JP"]], preds[["CN"]], preds[["ZA"]], preds[["SA"]], preds[["IR"]], preds[["TZ"]], preds[["ZM"]])
  colnames(out) <- c("predGDP")
  print(out)
  return(out)
}

@@

<<CD Lattice Graph>>=
usModel <- cobbDouglasModel("US")
coefs <- coef(usModel)
# print(coef(usModel))
usPred <- predict(usModel) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
class(usPred)
#print(usPred)
data.frame(usPred)

allData <- loadData("All")
predictions <- cobbDouglasPredictionsColumn()
allData <- cbind(allData, predictions)

print(allData)
@

>>>>

# Statistical significance levels. We'll work with 95% CIs
ciLevel <- 0.95
ciHalfLevel <- ciLevel + (1.0-ciLevel)/2.0
# List of countries
countries <- c(US="US", UK="UK", JP="JP", CN="CN", ZA = "ZA", SA="SA", IR="IR", TZ="TZ", ZM="ZM")
# Set the order for presenting countries in 3x3 lattice graphs. Default is alphabetical. "1" means first alphabetically.
# So, this list sets the order as US, UK, JP, CN, ZA, TZ, SA, IR, ZM.
countryOrderForGraphs <- c(7,6,3,1,8,5,4,2,9)
# Full page lattice plot sizes and other parameters
layoutSpec <- c(3,3) #indicates a 3x3 arrangement of panels.
figWidth <- 6.5 #Inches
figHeight <- 7.5 #Inches
keyTextSize <- 0.85 #85% of normal size
keyColumns <- 1 #Want only 1 column in the key for lattice graphs
keyXLoc <- 0.01 #x position of the key
keyYLoc <- 0.95 #y position of the key
scaleTextSize <- 0.75 #75% of normal size
scaleTickSize <- -0.5 #50% of normal size and pointing INWARD!
# List of energy types
energyTypes <- c(Q = "Q", X = "X", U = "U")
@

@@

<<Country Factors Graph, echo=FALSE>>=
createCountryFactorsGraph <- function(countryName){
  dataTable <- loadData(countryName)
  graphType <- "l"
  lineTypes <- c(1, 5, 2, 4, 6) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(1, 1, 1, 1, 1) #line widths
  colors <- c("black", "black", "black", "black", "black") #line and point colors
  lineSpec <- list(lty=lineTypes, lwd=lineWidths, col=colors)
  graph <- xyplot(iCapStk+iLabor+iQ+iX+iU ~ Year, data=dataTable,
                  type=graphType,
                  par.settings = list(superpose.line = lineSpec),
                  key=list(text=list(c("k", "l", "q", "x", "u")),
                           type=graphType,
                           lines=lineSpec,
                           columns=1, x=0.0, y=0.98),
                  scales=list(cex=1.0, #controls text size on scales
                              tck=scaleTickSize), #controls tick mark length
                  ylab="Indexed (1980=1)")
  return(graph)
}
@
<<Factors_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="GDP ($y$), capital stock ($k$), labor ($l$), thermal energy ($q$), exergy ($x$), and useful work ($u$) for all economies. (China's indexed GDP and indexed capital stock rise to $y = 7.3$ and $k = 9.2$ in 2011.)", echo=FALSE>>=
createFactorsLatticeGraph <- function(){
  dataTable <- loadData("All")
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1, 5, 2, 4, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 2, 2, 1, 1, 1) #line widths. 0 means no line.
  colors <- c("black", "black", "black", "red", "blue", "black") #line and symbol colors
  symbols <- c(1, NA, NA, NA, NA, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+iCapStk+iLabor+iQ+iX+iU ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, 
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "k", "l", "q", "x", "u")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createFactorsLatticeGraph()
@

@@

#########################
# Takes the Cobb-Douglas fitted models and creates a single column of predicted GDP values
# that corresponds, row for row, with the AllData.txt file.
##
cobbDouglasPredictionsColumn <- function(){
  preds <- lapply(countries, cobbDouglasPredictions)
  out <- rbind(preds[["US"]], preds[["UK"]], preds[["JP"]], preds[["CN"]], preds[["ZA"]], preds[["SA"]], preds[["IR"]], preds[["TZ"]], preds[["ZM"]])
  colnames(out) <- c("predGDP")
  print(out)
  return(out)
}

@@

<<US_CD_Graph>>=
usModel <- cobbDouglasModel("US")
# print(coef(usModel))
usPred <- predict(usModel) #See http://stackoverflow.com/questions/9918807/how-get-plot-from-nls-in-r
#print(usPred)
data.frame(usPred)

# allData <- loadData("All")
# predictions <- cobbDouglasPredictionsColumn()
# allData <- cbind(allData, predictions)
# 
# print(allData)
@

<<CD_Lattice_Graph, fig.width=figWidth, fig.height=figHeight, fig.cap="Cobb-Douglas model results.", echo=FALSE>>=
createCDLatticeGraph <- function(){
  dataTable <- loadData("All") #Grab the raw data
  predictions <- cobbDouglasPredictionsColumn() #Get the predictions from the models
  dataTable <- cbind(dataTable, predictions) #Now add the predictions column to the data. Column name = "predGDP".
  graphType <- "b" #b is for both line and symbol
  lineTypes <- c(0, 1) #line types. See http://en.wikibooks.org/wiki/R_Programming/Graphics
  lineWidths <- c(0, 1) #line widths. 0 means no line.
  colors <- c("black", "black") #line and symbol colors
  symbols <- c(1, NA)  #NA gives no symbol.
  graph <- xyplot(iGDP+predGDP ~ Year | Country, data=dataTable,
                  type=graphType,
                  index.cond=list(countryOrderForGraphs), #orders the panels.
                  layout=layoutSpec, #indicates a 3x3 arrangement of panels.
                  as.table=TRUE, #indexing of panels starts in upper left and goes across rows.
                  lty = lineTypes, lwd = lineWidths, col = colors, #Controls line parameters
                  pch = symbols, col.symbol = colors, #Controls symbol parameters
                  key=list(text=list(c("y", "y_pred")),
                           type=graphType,
                           cex=keyTextSize, #controls size of text in the key
                           lines=list(lty=lineTypes, lwd=lineWidths), #controls line types
                           pch=symbols, col=colors, #controls symbol (plot characters) types
                           columns=keyColumns, x=keyXLoc, y=keyYLoc), #controls columns and position of the key
                  scales=list(cex=scaleTextSize, #controls text size on scales.
                              tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                              y=list(at=c(1,2,3,4,5,6,7,8,9,10))), #tick mark locations
                  ylim=c(0.5,4.5), #y axis limits
                  xlab="", ylab="Indexed (1980=1 or 1991=1)") #axis labels
  return(graph)
}
createCDLatticeGraph()
@
